---
title: "PS-GC-LN-05-Raster calculations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)
require(knitr)
opts_knit$set(root.dir = '/home/hanna/Documents/Lehre/2016/Projektseminar/Excercises/session05/')
```

### Load libraries und set data path
```{r message=FALSE}
library(rgdal)
library(raster)

setwd("/home/hanna/Documents/Lehre/2016/Projektseminar/Excercises/session05/")

```

### 1) Simple raster calculations

The WorldClim Data are usually scaled into °C*10 (see http://www.worldclim.org/formats). 
Change the unit of the June temperature raster to °C instead of °C*10
```{r }
t6 <- raster("tmean_deu/tmean_deu_m06.tif")
t6 <- t6/10
```

Calculate the average temperatures of the summer months (June to August)

```{r }

t7 <- raster("tmean_deu/tmean_deu_m07.tif")
t8 <- raster("tmean_deu/tmean_deu_m08.tif")

summer_t <- (t6+t7/10+t8/10)/3
#or
summer_t <- mean(t6,t7/10,t8/10)
```

### 2) Calculations with RasterStacks
Calculate the average yearly temperatures
```{r}
files <- list.files("tmean_deu",pattern=".tif$",full.names=T)
yeart <- stack(files) #stack the indivisual files
yeart
yeart <- yeart/10 #divide all by 10
yearly_mean <- calc(yeart,mean) #calculate the average
#or 
yearly_mean <- mean(yeart) #calculate the average
spplot(yearly_mean)
yearly_sd <- calc(yeart,sd) #calculate the variability
spplot(yearly_sd)
```
Calculate one single average value of the yearly temperatures in Germany (use ?values to extract the values of a raster)
```{r }
meanValue <- mean(values(yearly_mean),na.rm=TRUE)
#or
meanValue <- cellStats(yearly_mean,mean)
meanValue
```
### 3) Mask rasters

Use the gadm data to mask your raster of average yearly temperatures. 
Then re-calculate your single average value of the yearly temperatures in Germany.

```{r}
lyr <- ogrListLayers("DEU_adm1.shp")[1]
gadm <- readOGR("DEU_adm1.shp",layer=lyr)
yearly_mean_DEU <- mask(yearly_mean,gadm)
spplot(yearly_mean_DEU)

meanValue <- mean(values(yearly_mean_DEU),na.rm=TRUE)
meanValue
```


### 4) Reclassifications
Create a raster that distinguish between areas with below average yearly temperatures and above average yearly temperatures based on the single average value of the yearly temperatures in Germany.

```{r}
rcl <- matrix(c(-999,meanValue,1,
                meanValue,999,2),
              ncol=3, byrow = TRUE)
rcl
classified <- reclassify(yearly_mean_DEU, rcl)
spplot(classified)
```
