---
title: "PS-GC-LN-03-Projektionen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(knitr)
opts_knit$set(root.dir = '/home/hanna/Documents/Lehre/2016/Projektseminar/Excercises/session2/')
```

### Load libraries und data
```{r}
library(raster)
library(rgdal)

setwd("/home/hanna/Documents/Lehre/2016/Projektseminar/Excercises/session2/")

tmean <- raster("tmean_deu_m01.tif")
lyr <- ogrListLayers("gadm_deu.shp")[1]
gadm <- readOGR("gadm_deu.shp",layer=lyr)
lyr <- ogrListLayers("Hauptstrassen.shp")[1]
roads <- readOGR("Hauptstrassen.shp",layer=lyr)
```

### 1) On the fly Projection?

Plot tmean. Overlay the gadm data. What happens? Next, add the roads to the plot.
Is there a problem?

```{r }
plot(tmean)
plot(gadm,add=T)
plot(roads,add=T)
```

### 2) Show information of projections
```{r }
tmean
proj4string(tmean)
gadm
proj4string(gadm)
proj4string(roads)
```

### 3) Reproject raster und vector data

Project tmean and gadm into UTM. Then plot tmean and add gadm and roads.
Still a problem?

```{r }
tmean_utm <- projectRaster(tmean, crs="+proj=utm +zone=32 +ellps=WGS84 +datum=WGS84 +units=m +no_defs" )
```
oder übernehme einen proj4String:
```{r }
tmean_utm <- projectRaster(tmean, crs=proj4string(roads))
```
Das gleiche für Vektordaten:
```{r }
gadm_utm <- spTransform(gadm,proj4string(roads))
```
Plotten der Ergebnisse
```{r }
plot(tmean_utm)
plot(gadm_utm,add=T)
plot(roads,add=T,col="red")
```


### 4) Assign/Define projections

Assign the correct projection to a new raster.

```{r }
tmean_02 <- raster("tmean_deu_m02_noProj.tif")
proj4string(tmean_02) 
tmean_02 # Koordinaten sind geokoordinaten also ist latlon wgs 84 naheliegend
proj4string(tmean_02) <- "+proj=longlat +ellps=WGS84 +towgs84=0,0,0,0,0,0,0 +no_defs "
tmean_02
```
